---
title: "Practical Machine Learning"
author: "Gautam"
date: "25 October 2015"
output: html_document
---

# Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

## Loading libraries and setting seed

```{r}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(randomForest)
library(knitr)
set.seed(13433)
```

Now this is done, next we read data from Url and send it to train and test   

```{r}
trainUrl<-"http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl<-"http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
```
Giving the Url to a variable and removing all the NA and useless values   
```{r}
train<-read.csv(url(trainUrl), na.strings=c("NA","","#DIV/0!"))
test <-read.csv(url(testUrl), na.strings=c("NA","","#DIV/0!"))

```

## Getting and cleaning data

We have to only take the parameters relevant to the analysis. Hence this step is done.  

```{r}
missingData = is.na(train)
sensorColumns = grep(pattern = "_belt|_arm|_dumbbell|_forearm", names(train))
length(sensorColumns)
data = train[, c(sensorColumns,160)]
dim(data)
omitColumns = which(colSums(missingData) > 19000)
data = data[, -omitColumns]
dim(data)
```

## Partioning the data into training and testing
I have split the data into 60% training and 40% testing
```{r}
intrain=createDataPartition(train$classe,p=0.6,list=FALSE)
trainset<-train[intrain,]
testset<-train[-intrain,]

```



## Removing nearZeroVariances

Run this code to create another subset without NZV variables:  

```{r}
myDataNZV <- nearZeroVar(trainset, saveMetrics=TRUE)


myNZVvars <- names(trainset) %in% c("new_window", "kurtosis_roll_belt", "kurtosis_picth_belt",
                                      "kurtosis_yaw_belt", "skewness_roll_belt", "skewness_roll_belt.1", "skewness_yaw_belt",
                                      "max_yaw_belt", "min_yaw_belt", "amplitude_yaw_belt", "avg_roll_arm", "stddev_roll_arm",
                                      "var_roll_arm", "avg_pitch_arm", "stddev_pitch_arm", "var_pitch_arm", "avg_yaw_arm",
                                      "stddev_yaw_arm", "var_yaw_arm", "kurtosis_roll_arm", "kurtosis_picth_arm",
                                      "kurtosis_yaw_arm", "skewness_roll_arm", "skewness_pitch_arm", "skewness_yaw_arm",
                                      "max_roll_arm", "min_roll_arm", "min_pitch_arm", "amplitude_roll_arm", "amplitude_pitch_arm",
                                      "kurtosis_roll_dumbbell", "kurtosis_picth_dumbbell", "kurtosis_yaw_dumbbell", "skewness_roll_dumbbell",
                                      "skewness_pitch_dumbbell", "skewness_yaw_dumbbell", "max_yaw_dumbbell", "min_yaw_dumbbell",
                                      "amplitude_yaw_dumbbell", "kurtosis_roll_forearm", "kurtosis_picth_forearm", "kurtosis_yaw_forearm",
                                      "skewness_roll_forearm", "skewness_pitch_forearm", "skewness_yaw_forearm", "max_roll_forearm",
                                      "max_yaw_forearm", "min_roll_forearm", "min_yaw_forearm", "amplitude_roll_forearm",
                                      "amplitude_yaw_forearm", "avg_roll_forearm", "stddev_roll_forearm", "var_roll_forearm",
                                      "avg_pitch_forearm", "stddev_pitch_forearm", "var_pitch_forearm", "avg_yaw_forearm",
                                      "stddev_yaw_forearm", "var_yaw_forearm")
trainset <- trainset[!myNZVvars]
```

## Eliminating the columns with large number of NA values
```{r}
trainset <- trainset[c(-1)]
trainingV3 <- trainset #creating another subset to iterate in loop
for(i in 1:length(trainset)) { 
  if( sum( is.na( trainset[, i] ) ) /nrow(trainset) >= .6 ) { #if n?? NAs > 60% of total observations
    for(j in 1:length(trainingV3)) {
      if( length( grep(names(trainset[i]), names(trainingV3)[j]) ) ==1)  { #if the columns are the same:
        trainingV3 <- trainingV3[ , -j] #Remove that column
      }   
    } 
  }
}
```

## Trimming the trainset and testset

```{r}

trainset <- trainingV3
rm(trainingV3)
clean1 <- colnames(trainset)
clean2 <- colnames(trainset[, -58]) #already with classe column removed
testset <- testset[clean1]
test <- test[clean2]

for (i in 1:length(testset) ) {
  for(j in 1:length(trainset)) {
    if( length( grep(names(trainset[i]), names(testset)[j]) ) ==1)  {
      class(testset[j]) <- class(trainset[i])
    }      
  }      
}

#And to make sure Coertion really worked, simple smart ass technique:
test <- rbind(trainset[2, -58] , test) #note row 2 does not mean anything, this will be removed right.. now:
test <- test[-1,]

```

## Predicting with decision trees

```{r}
modFitA1 <- rpart(classe ~ ., data=trainset, method="class")
p<-predict(modFitA1,newdata=testset,type="class")
confusionMatrix(p,testset$classe)
```

## Predicting with Random Forests

```{r}
modFitB1 <- randomForest(classe ~. , data=trainset)
predictionsB1 <- predict(modFitB1, testset, type = "class")
confusionMatrix(predictionsB1, testset$classe)

```






#Conclusion
It can be seen that Random Forests give better prediction than Decision trees.